---
title: "Stats_data_rich"
output: html_document
date: "2026-09-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Review of Hypothesis Testing

Null - the default state of the system which we are testing; denoted $H_0$

Alternative - not the null; denoted $H_1$

Significance level - defines threshold for Type I errors; denoted $\alpha$ 

p-value - probability of obtaining a result as extreme as or more extreme than the data under the null; $\mathbb{P}(\text{data as or more extreme } | H_0 )$

```{r}
# set a random seed for reproducibility
set.seed(222)
p <- 0.5 # prob. of head for fair coin
flips <- 1000
data <- sample(c('H', 'T'), 
               flips, prob = c(p, 1-p), 
               replace = TRUE)
heads <- sum(data == 'H')

# or, for a faster vectorized version
heads <- rbinom(1, flips, p)

# compute a one sided p-value 
one.sided.pvalue <- 1 - pbinom(heads, flips, 0.5)
print(one.sided.pvalue)

# and compare to a two sided p-value
heads <- rbinom(1, flips, p)
two.sided.pvalue <- binom.test(heads, flips, 0.5, alternative = 'two.sided')
print(two.sided.pvalue)

# plot a distribution of flips
heads_dist <- rbinom(1000, flips, p)
hist(heads_dist, main='distribution number heads', xlab='number of heads')
```

```{r}
# function to compute the distribution of p-values
plot_pvalue_dist <- function(num_coins, flips, p){
  heads <- rbinom(num_coins, flips, p)
  pvalues <- 1-pbinom(heads, flips, 0.5)
  hist(pvalues, xlab='p-value', freq = FALSE)
  abline(h=1, col='red', main = paste0('p = ', p), lty=2)
}
# simulate
num_coins <- 25000
plot_pvalue_dist(num_coins, flips, p)

# compare that to a biased coin
p <- 0.52
plot_pvalue_dist(num_coins, flips, p)

p <- 0.55
plot_pvalue_dist(num_coins, flips, p)


```

## p-hacking

- Data trimming (excluding outliers)
- Picking different variables and fishing for significance
- Data transformation
- Stopping data collection early
- Hypothesis switching
- Data mining
- Selecting test statistics ('p-value shopping')
- Overfitting (fitting complex models with additional parameters until significance is achieved... e.g. adding regressors to a multiple regression until the coefficient for a given variable is significant)

### Type I and II Errors

- Type I - false positive; test is significant and $H_0$ is true; controlled by $\alpha$
- Type II - false negative; test is not significant and $H_0$ is false; sensitive to power and sample size

### Distribution of p-values
- Uniform if the null is true; plot with histogram or Q-Q plot

## Multiple Testing
- Let S denote true positives, V denote false positives (Type I errors), T denote false negatives (Type II errors), and U denote true negatives

- FWER $ = 1 - \mathbb{P}(V = 0) = \mathbb{P}(V \geq 1) = 1 - (1- \alpha)^{n}$ where $n$ is the number of tests; probability of at least one Type I error occurring

- FDR = $\mathbb{E}[V / (V+S)]$ is the expected proportion of positive tests which are false; most often we will define a new significance level $\alpha^* < \alpha$ s.t. $FDR < \alpha$ where $\alpha$ is the original or desired significance level

- Bonferroni Correction - adjust p-values by the number of tests ($n$); concretely, we set $p_i^* = \min(p_i*n, 1)$

- Benjamini-Hochberg Correction - order the p-values and set $p_i = \min(\frac{i*p_i}{n}, 1)$ and reject up to $p_i < \alpha$; equivalently, find the smallest $i$ s.t. $\frac{i*p_i}{n} < \alpha$

```{r}
toss_coins <- function(num_flips, p){
  heads <- rbinom(1, num_flips, p)
  pvalue <- 1 - pbinom(heads, num_flips, 0.5)
  if (p == 0.5){
    return(data.frame(p=p, heads=heads, pval=pvalue, H0=TRUE))
  } else {
    return(data.frame(p=p, heads=heads, pval=pvalue, H0=FALSE))
  }
}

set.seed(8)
data <- data.frame()
# biased samples
for (i in 1:50) data <- rbind(data, toss_coins(1000, 0.55))
# unbiased samples
for (i in 1:50) data <- rbind(data, toss_coins(1000, 0.5))
head(data)

library(dplyr)
data <- arrange(data, pval)
head(data)
```

## Regression and Model Selection

### Linear Regression

- Model - $\hat{Y} = mX + b + \epsilon$ and $Y$ is the observed data; to find the OLS solution, we solve $\arg \min_{m, b} (\hat{Y} - Y)^2$

- $R^2 = \text{var}(\hat{Y}) / \text{var}(Y)$ 

### Logistic Regression

- Model - $Y \in \{0, 1\}$ (this could be disease state or some other discrete RV)
- Convert $Y$ to continuous... consider the probability of a disease, $\mathbb{P}(\text{disease}) \in [0,1]$ and then take the odds ratio, $OR = \frac{\mathbb{P}(\text{disease})}{1-\mathbb{P}(\text{disease})} \in [0, \infty)$ and finally compute the logarithm (i.e., the logit function) as $\log(OR) \in (-\infty, \infty)$

### AIC and BIC

### Cross Validation